Paper without a name

### 2014-05-23 Article.tex
Changes...

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.0in]{geometry}
\usepackage[british]{babel}
\usepackage{indentfirst}
\usepackage{array,booktabs,longtable}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{url}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

@@

@@

Mathematically, we consider that the wires are numbers from 1 to $n$. A box is a
permutation $\tau(\epsilon) = (i,j)^{\epsilon}$, where $\epsilon$ is equal to 0
or 1.  A configuration of boxes C is the concatenation of k boxes $\tau_1,
\ldots, \tau_k$, that means, $C(\epsilon_1, \ldots, \epsilon_k) =
\tau_1^{\epsilon_1}\circ\ldots\circ\tau_k^{\epsilon_k}$.  The configuration C is
said valid if $C(\{0,1\}^k)= \mathfrak{S}_n$.

@@

We can easily determine trivial lower and upper bounds of the number of
boxes~:\\ Given the fact that C is an application, we need to have
$Card(\{0;1\}^{k}) \geq Card(\mathfrak{S}_n)$. We immediateley obtain that $k$
has to be greater than $\log_2(n!)$.\\ We also have an upper bound~: The
traditionnal bubblesort is a valid configuration of the SwitchBoxes problem. We
obtain that the optimal k is smaller than $\frac{(n-1)(n-2)}{2}$.

In this contribution, we first develop how the optimal k seems to be linked to
the maximal number of comparisons for sorting n elements by binary insertion.
We explain why we guess that those numbers are equal.

@@

% Part 0. Introduction
% Part 1. Conjecture
% Part 2. Optimisation
% Part 3. Pistes de recherche

@@

Using this algorithm, we got the following results~:

\begin{tabular}{|c|c|}
\hline
Number of wires $n$ & $opt_n$ \\
\hline
2 & 1 \\
3 & 3 \\
4 & 5 \\
5 & 8 \\
6 & 11 \\
\hline
\end{tabular}\\

According to the fact that this sequence is exactly the beginning of the
sequence A001855 of the Encyclopedia of Integer Sequences\footnote{The On-Line
Encyclopedia of Integer Sequences, Sorting numbers: maximal number of
comparisons for sorting n elements by binary insertion,
\url{http://oeis.org/A001855}}, corresponding to the maximal number of
comparisons for sorting $n$ elements by binary insertion, we guessed that these
sequences were equal.

@@

$n=4$ : fill in later

@@

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.0in]{geometry}
\usepackage[british]{babel}
\usepackage{indentfirst}
\usepackage{array,booktabs,longtable}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{url}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathtools}

@@

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

@@

Mathematically, we consider that the wires are numbers from 1 to $n$. A box is a
permutation $\tau(\epsilon) = (i,j)^{\epsilon}$, where $\epsilon$ is equal to 0
or 1.  A configuration of boxes $C$ is the concatenation of $k$ boxes $\tau_1,
\ldots, \tau_k$, that means, $C(\epsilon_1, \ldots, \epsilon_k) =
\tau_1^{\epsilon_1}\circ\ldots\circ\tau_k^{\epsilon_k}$.  The configuration $C$ is
said valid if $C(\{0,1\}^k)= \mathfrak{S}_n$.

@@

We can easily determine trivial lower and upper bounds of the number of
boxes~:\\ Given the fact that $C$ is an application, we need to have
$Card(\{0;1\}^{k}) \geq Card(\mathfrak{S}_n)$. We immediateley obtain that $k$
has to be greater than $\log_2(n!)$.\\ We also have an upper bound~: The
traditionnal bubblesort is a valid configuration of the SwitchBoxes problem. We
obtain that the optimal $k$ is smaller than $\frac{(n-1)(n-2)}{2}$.

In this contribution, we first recall how the sequence of the maximal number of
comparisons for sorting $n$ elements by binary insertion is defined and develop
how the optimal $k$ seems to be linked to it.  We then explain why we guess that
those numbers are equal.

@@

\section{The A001855 sequence}

In this paragraph, we will develop on the A001855 sequence, here denoted $u_n$,
which seems to be in relation with our problem.

$u_n$ is defined as the maximal number of comparisons done by a binary
insertion sort running on a table of $n$ items.

\begin{verbatim}

    function FindPos(n, T, begin, end)   // find n in T(begin, begin+1, ..., end-1)
        assert(begin < end)
        if begin + 1 = end then
            return begin
        else
            mid = E[(begin + end) / 2]
            if T[mid] > n then
                return FindPos(n, T, begin, mid)
            else
                return FindPos(n, T, mid+1, end)

    function Sort(T)
        n = T.size
        S = [T[0]]
        for i = 1 to n-1
            pos = FindPos(T[i], S, 0, S.size)
            insert T[i] in S at position pos
        return S

\end{verbatim}

This algorithm progressively builds $S$, the table of the elements of $T$ in
increasing order. This is done by progressively inserting the elements of $T$ in
$S$ at a position calculated by the $FindPos$ function. The $FindPos$ function
is based on a binary search, therefore does $O(log(S.size))$ comparisons.

The $u_n$ suite is defined as follows :

$$u_1 = 0$$
$$\forall n \in \mathbb{N}^*, u_{n+1} = u_n + \ceil{log_2(n)}$$

This sequence is described on the OEIS\footnote{Online Encyclopedia of Integer
Sequence}, and is referenced under the name A001855.

@@

Using this algorithm, we got the results in the table of figure~\ref{box-count}.

\begin{figure}
    \centering
    \begin{tabular}{|c|c|}
    \hline
    Number of wires $n$ & $opt_n$ \\
    \hline
    2 & 1 \\
    3 & 3 \\
    4 & 5 \\
    5 & 8 \\
    6 & 11 \\
    \hline
    \end{tabular}
    \caption{Minimum box count for small number of wires}
    \label{box-count}
\end{figure}

According to the fact that this sequence is exactly the beginning of the $u_n$
sequence described in the previous paragraph, our conjecture is that the
sequences are equal.

@@

$n=4$ : $S_4(e_1,e_2,e_3,e_4,e_5) = (1, 2)^{e_1} \circ (3,4)^{e_2} \circ
(1,3)^{e_3} \circ (2, 4)^{e_4} \circ (1,2)^{e_5}$

@@

For a box system $S = b_1 b_2 \dots b_p$, we construct the sets :

$$Perm_0 = { id }$$
$$Perm_k = (b_1 b_2 \dots b_k)(\left\{0,1\right\}^k)$$

The permutation set $Perm_p$ therefore contains all the possible permutations we
can generate with the $p$ boxes. If $Perm_p = \mathfrak{S}_n$ then the box system
is valid.

Each set $Perm_k$ is represented in memory as a bitset containing $n!$ bits, one
for each permutation of $\mathfrak{S}_n$. The isomorphism between the
permutations of $\mathfrak{S}_n$ and the numbers $\left\{0,1,\dots,n!-1\right\}$ is
constructed using the standard method described in (a citation).

The representation as a bitset is the most compact representation we have found.

@@

We can easily detect that $Perm_{k+1} = Perm_k$ when $ \# Perm_{k+1} = \# Perm_k$,
because $Perm_k \subset Perm_{k+1}$. When this condition is fulfilled, then we
know that the box $b_{k+1}$ is useless because it adds no permutation to the set
of generated permutations. We can therefore remove it from the box system.

By removing the boxes that this method detects as useless from the box system we
constructed in the previous paragraphs, we have constructed box systems that
generate all the permutations for $n \leq 13$ and that have $u_n$ boxes (ie the
conjectured optimum). We have not managed to run the test for $n \geq 14$,
because the required RAM exceeds 20Go, which is not a resource we have in our
possession.
